### 2017-12-07 paper-proposal.tex
Good job. I think we should focus on the "big and small problem" in the introduction, then lead into "in this study ..."  and how we solved the problems, and finally conclude with the benefits spfy offers.

\bibliography{paper-webserver.bib}

\begin{document}
%contain, at the top, the following affirmative statement. "This website is free and open to all users and there is no login requirement." Additionally, any third party software employed by the website that has more restrictive usage terms must be listed.
This website is free and open to all users and there is no login requirement. The code for this webserver and all third party software used by this website are available under the open-source Apache 2.0, BSD 3-clause, or similar licenses.

% include the website address; website name; and the names, affiliations, and email addresses of all authors.
The website is available at \url{https://lfz.corefacility.ca/superphy/spfy/}. Spfy's codebase is provided at \url{https://github.com/superphy/backend} and the documentation at \url{https://superphy.readthedocs.io/en/latest/}.

% include a notification if this is an update from a previous publication in the Web Server issue, and in that case, include an estimate of the number of users and the number of citations.
% For web servers, or essentially similar web servers, that have been the subject of a previous publication, including publication in journals other than NAR, there is a minimum two-year interval before re-publication in the Web Server Issue.
Our proposal covers an update to Superphy \citep{whiteside2016superphy}, an online predictive genomics platform targeting \textit{Escherichia coli}.
The update, called Spfy, uses graph datastructures to store and retrieve results for computational workflows.
We demonstrate the ability of graph data structures to scale to the [approximate number, eg. greater than 50,000]l of whole-genome sequences  accumulated so far, and show the ability to scale to X genomes.

Spfy provides real-time subtyping through a ReactJS-based website, where user-defined analyses options are selected. Analyses are run in parallel using task queues and Docker containerization, and the results are immediately displayed to the user following their completion.
Subtyping options include O-antigen, H-antigen, Shiga-toxin 1 (Stx1), Shiga-toxin 2 (Stx2), and Intimin typing. Reference-lab type tests include virulence factor and anti-microbial resistance annotation. All genomes are analyzed withing the pan-genome framework of \textit{E. coli}.

% IF THE WEBSITE IMPLEMENTS A META-SERVER OR COMPUTATIONAL WORKFLOW, the summary MUST describe 1) significant added value beyond the simple chaining together of existing third party software or the calculation of a consensus prediction from third party predictors and classifiers; and at least one of the following: 2) how user time for data gathering and multi-step analysis is significantly reduced, or 3) how the website offers significantly enhanced display of the data and results.
Current comparative computational workflows chain different analysis software, but lack storage and retrieval methods for generated results.

Within Spfy, we store the output from every analysis, and link them together in the context of a genome graph. This graph also stores metadata for each genome, facilitating inquiries ranging from population genomics to epidemiological investigations.

% provide descriptions of the input data, the output, and the processing method; complete citations for previous publications of the method or the web server; and two to four keywords. Additionally, authors must indicate how long the server has been running, the number of inputs analyzed during testing, and an estimate of the number of individuals outside of the authors' group who have been involved in the testing.
Spfy currently contains 59,5323 public \textit{E. coli} assembled genomes, 5,353 genomes from GenBank and 54,181 genomes from Enterobase (\~596 GB), storing the genome sequences and results from all analyses.

The resulting database had XYZ million nodes and XYZ million edges, with XYZ object properties, which worked out to XYZ TB of data stored.
% up-time

% collaborators

% existing
Existing scientific workflow technologies such as Galaxy \citep{goecks2010galaxy}, and pipelines such as the Integrated Rapid Infectious Disease Analysis (IRIDA) \url{http://www.irida.ca/} and the Bacterium Analysis Pipeline (BAP) \citep{thomsen2016bacterial} are well established.
% data integration
Results from these workflows are stored in tables, or the file is saved on disk.
Because output from different analyses are structured differently using distinct terminology, formats must be converted before they can be compared. Without a unified structure, these conversions quickly become impractical for broad usage.
% ontologies
To avoid proliferating ontologies, and to allow Spfy to integrate with existing ones, annotations from the GenEpiO \citep{griffiths2017context}, FALDO \citep{bolleman2016faldo}, and TypOn \citep{vaz2014typon} ontologies are used to describe biological data.

% analysis run-time / throughput with different levels of parallelization