### 2018-03-01 tables/comparison.tex
ADD: touching on patric

% Please add the following required packages to your document preamble:
% \usepackage{graphicx}
\begin{table}[!hb]
\centering
\caption{Comparison of various bioinformatic pipelines and their underlying database. Functionally, Spfy integrates different analysis modules as in BAP while also merging large datasets as in PATRIC.}
\label{Table 2.}
\resizebox{\textwidth}{!}{%
\begin{tabular}{lllll}
         & Spfy                & Galaxy              & BAP                 & PATRIC              \\
Database & Blazegraph          & PostgreSQL          & MySQL + File System & MongoDB + Shock     \\
Type     & Graph               & SQL                 & SQL + File System   & NoSQL               \\
Focus    & Integrated Analyses & Workflow Technology & Batch Analysis      & Integrated Analyses
\end{tabular}%
}
\end{table}